---
  title: crosslink An R package for network visualization of grouped nodes
  output: html_document
---


The goal of crosslink is to visualize the network of grouped nodes

## 1. Installation

You can install the released version of crosslink from [github](https://github.com/zzwch/crosslink) with:

```{r}
#devtools::install_github("zzwch/crosslinks") 
#install.packages("aplot")
```

## 2. Quick start

Examples of typical crosslonk usage.

```{r}
#library(crosslink)

load("data/demo.RData")

cl <- crosslink(demo$nodes, demo$edges, demo$cross.by, odd.rm = F,spaces = "flank")
cl %>% cl_plot()

```

## 3. Step by step

This is a basic example of the basic function of crosslink packages,including 
1). Input data
2). Generate CrossLink class
3). Coordinate transformation
4). Layout modules
5). Plotting modules

### 1). Input data

crosslink needs two files as input. 
a. nodes (must have two columns: node name and node type)
b. edges data (must have two columns: source node and target node). 

crosslink also provides the function 'gen_demo' to generate demo data. 

```{r}
#library(crosslink)

# generate demo data
n <- 6
demo <- gen_demo(n_cross = n,n_node = 4:(n+3), n_link = 3:(n+1), seed = 66)
nodes <- demo$nodes
edges <- demo$edges
cross.by <- demo$cross.by

```

### 2). Generate crossLink class

crosslink can help to generate an object of crosslink class for plot.


```{r}
# users can define 'odd.rm' to choose if remove the nodes have zero relationship with any other nodes when generate crosslink class
# user can set intervals between nodes and gaps through spaces and gaps.
cl <- crosslink(nodes, edges, cross.by, odd.rm = F,spaces = "flank")

cl %>% get_cross()   # get node information
cl %>% get_link()    # get edges information
cl %>% cl_layouts()  # get layouts information of crosslink object
cl %>% cl_active()   # get active layouts information of crosslink object

```


### 3). Coordinate transformation

Coordinate transformation consists of affine transformation and functional transformation. 
The `tf_affine` function contains `tf_rotate`, `tf_shift`, `tf_shear`, `tf_flip` and `tf_scale` function. 
The `tf_fun` interface allows user to custom transforming function.

```{r}

# tf_rotate, rotating in a specific angle with (x,y) as the center.
cl %>% tf_rotate(x=0,y=1,angle = 45) %>% cl_plot()

# tf_shift, shifting a relative distance according to x-axis or y-axis
cl %>% tf_shift(x=1,y=-1) %>% cl_plot()

# tf_shear
cl %>% tf_shear(axis = "x",angle = 60) %>% cl_plot()

# tf_flip, flip the figure according to x-axis or y-axis
cl %>% tf_flip(axis = "y") %>% cl_plot()

# tf_scale
cl %>% tf_scale(x=0,y=1,scale = 5) %>% cl_plot() 

# tf_fun, coordinate transformation according to custom-defined function
cl %>% tf_fun(fun = sin,along = "y",xrange.from=c(0,0.5*pi)) %>% cl_plot()

# combined tf functions
cl %>% tf_flip(axis = "y") %>% tf_fun(fun = sin,along = "y",xrange.from=c(0,0.5*pi)) %>% cl_plot()

```

### 4). Layout modules


a. Commonly used styles are predefined in layout module, including row, column, arc, polygon and hive.
crosses can be specified in all five layout module
b. ‘set_header’ function is provided to conveniently place cross headers

```{r}
# default layout module is column
cl %>% cl_plot()
cl %>% layout_column() %>% cl_plot()

# layout by row
cl %>% layout_row() %>% cl_plot()

# layout by arc
cl %>% layout_arc(angles = 60,crosses = c("E","F")) %>% cl_plot()

# layout by polygon (list of angles must have the same length with crosses)
cl %>% layout_polygon() %>% cl_plot()

# layout by hive
cl %>% layout_hive() %>% cl_plot()

# Header can be customized through 'set_header'
set_header(cl,header = c("A","B","C","D","E","F")) -> cl
cl %>% get_header()  # get header of crosslink object 

```


### 5). Plotting modules

We introduce this pattern in three aspects. 

a. quick plotting (`cl_plot`)
b. aesthetic settings ( The color, size, type and text of the nodes and lines in the network)
c. combination of the network diagram with the corresponding node annotation image in aligning coordinates

#### a. quick plotting
```{r}
cl %>% cl_plot()
```


#### b. aesthetic settings
```{r}
# aesthetic settings based on ggplot2 system, some specific examples are shown below.

# set colors, shapes and size of nodes
# cross:a named list of arguments for crosses. usage same as ggplot2::geom_point(). Set NULL to use default settings, or Set NA to not show.
cl %>% cl_plot(cross = list(mapping =  aes(color = type,shape= type,size=type),
                            scale   = list(color = scale_color_manual(values = RColorBrewer::brewer.pal(8, "Set1")),
                                           size  = scale_size_manual(values = seq(1,12,by=2)),
                                           shape  = scale_shape_manual(values = c(15:22))
                                          )
                           ))

                           
# set colors, linetypes and size of edges
# link: a named list of arguments for links. usage same as ggplot2::geom_segment(). Set NULL to use default settings, or Set NA to not show.
cl %>% cl_plot( link = list(mapping =  aes(color = src.cross,linetype =src.cross),
                            scale   = list(color = scale_color_manual(values = RColorBrewer::brewer.pal(8, "Set1")),
                                           linetype =scale_linetype_manual(values = c(1:6))),
                            size    = 1.5
                           ))
                            
# set header styles
# header:	a named list of arguments for headers. usage same as ggplot2::geom_text(). Set NULL to use default settings, or Set NA to not show.
cl %>% cl_plot(header = list(mapping = aes(color= cross),
                               scale = list(color = scale_color_manual(values = RColorBrewer::brewer.pal(8, "Dark2"))),
                                size = 5
                             ))

# set figure theme
# add:	other gg object to be added to final plot, such as theme().
theme_use <- theme(legend.position = "none",aspect.ratio = 1,
                   axis.title = element_blank(),
                   axis.text = element_blank(),
                   axis.ticks = element_blank(),
                   panel.grid = element_blank(),
                   panel.background = element_blank())
cl %>% cl_plot(add=theme_use)

```


#### c. annotation figure
```{r}
# cl_annotation： add annotation figure. 
# top, bottom, left or  right	: ggplot object
# top.by, bottom.by, left.by, right.by : name of cross by which to align ggplot

ann.data <- data.frame(F=factor(paste0("F",c(1:6)),levels=paste0("F",c(6:1))),value=c(6:1))
ann.data %>% ggplot(mapping = aes(x=F,y=value))+geom_bar(stat="identity")+coord_flip() -> rgtAnn
cl %>% cl_plot(annotation=cl_annotation(right= rgtAnn,right.by ="F"))

```


## 4. Examples

There are examples used in paper and practical application.

### 1). examples used in paper

#### a. layout by row
```{r}
load("data/demo.RData")
cl <- crosslink(demo$nodes, demo$edges, demo$cross.by, odd.rm = F,spaces = "flank")

cl %>% layout_row() %>%
   cl_plot(cross = list(mapping = aes(color  = type,fill=type),
                        scale   = list(color = scale_color_manual(values = RColorBrewer::brewer.pal(8, "Dark2")),
                                        fill = scale_fill_manual( values = RColorBrewer::brewer.pal(8, "Dark2"))),
                        size=8,shape=24
                       ),
          link   = list(color=RColorBrewer::brewer.pal(8, "Set2")[7],size=1.5,linetype=12),
          label  = list(color="white"),
          header = list(mapping = aes(color= cross),
                        scale = list(color = scale_color_manual(values = RColorBrewer::brewer.pal(8, "Dark2"))),
                        size=5
                       ),
          add=theme_use)

```

#### b. layout by rotate
```{r}

cl %>% 
   tf_rotate(angle = 15, by.each.cross = T) %>%
   cl_plot(cross = list(mapping = aes(color = type),
                       scale = list(color = scale_color_manual(values = RColorBrewer::brewer.pal(8, "Dark2"))),
                       size=10,shape=16
                       ),
          link = list(color="grey75",size=1,linetype=15),
          label = list(color="white"),
          header = list(mapping = aes(color= cross),
                       scale = list(color = scale_color_manual(values = RColorBrewer::brewer.pal(8, "Dark2"))),
                       size=5
                       ),
          add=theme_use)
          
```

#### c. layout by polygon
```{r}
cl %>% layout_polygon() %>%
   #tf_rotate(angle = 45, by.each.cross = F) %>%  # If the grouping is 4, rotate by this parameter and change from diamond to square 
   cl_plot(cross = list(mapping = aes(color = type),
                        scale = list(color = scale_color_manual(values = RColorBrewer::brewer.pal(8, "Dark2"))),
                        size=10,shape=16
                       ),
           link  = list(color="orange",size=2,linetype=15,geom="curve"),
           label = list(color="white"),
           header= list(mapping = aes(color= cross),
                        scale = list(color = scale_color_manual(values = RColorBrewer::brewer.pal(8, "Dark2"))),
                        size=5
                        ),
           add  =  theme_use )

```

#### d. layout by hive

```{r}
cl %>% layout_hive(angles=rep(60,6)) %>% #length of angles must be same with the numbers of groups
       cl_plot(cross = list(mapping = aes(color = type),
                       scale = list(color = scale_color_manual(values = RColorBrewer::brewer.pal(8, "Dark2"))),
                       size=10,shape=18
                       ),
               link  = list(color="purple",size=1,linetype=15),
               label = list(color="white"),
               header= list(mapping = aes(color= cross),
                            scale = list(color = scale_color_manual(values = RColorBrewer::brewer.pal(8, "Dark2"))),
                            size=5
                            ),
               add   = theme_use )

```

#### e. layout by arc

```{r}
cl %>% layout_arc(angles=c(60,120,180),crosses=c("D","E","F")) %>% #length of angles must be same with the numbers of crosses
       cl_plot(cross = list(mapping = aes(color = type),
                            scale = list(color = scale_color_manual(values = RColorBrewer::brewer.pal(8, "Dark2"))),
                            size=10,shape=18
                           ),
               link  = list(color="orange",size=1,linetype=3),#,geom="curve"
               label = list(color="white"),
               header= list(mapping = aes(color= cross),
                            scale = list(color = scale_color_manual(values = RColorBrewer::brewer.pal(8, "Dark2"))),
                            size=5
                           ),
               add   =  theme_use)

```

#### f. layout by combination of several methods

```{r}

# demo data
n <- 5
demo <- gen_demo(n_cross = n, n_node = rep(n, n), n_link = rep(n, n-1), seed = 66)
nodes <- demo$nodes
edges <- demo$edges
cross.by <- demo$cross.by


cl <- crosslink(nodes, edges, cross.by, odd.rm = F)
cl <- set_header(cl,header = c("A","B","C","D","E"))

cl %>% 
   layout_polygon(crosses = c("A","B","C","D")) %>%
   tf_rotate(crosses=c("A","B","C","D"),angle = rep(45,4)) %>%
   tf_shift(x=0.8*(-1),y=0.05,crosses="E",layout="default") ->cl

cl %>%
   cl_plot(cross = list(mapping = aes(color = type),
                       scale = list(color = scale_color_manual(values = RColorBrewer::brewer.pal(8, "Dark2"))),
                       size=10,shape=19
                       ),
          link   = list(color="grey75",size=1,linetype=3),
          label  = list(color="white"),
          header = list(mapping = aes(color= cross),
                       scale = list(color = scale_color_manual(values = RColorBrewer::brewer.pal(8, "Dark2"))),
                       size=5
                       ),
           add   =  theme_use)
           
```

### 2). examples of complex figure

```{r}
source("R/data.R")
library(dplyr)
theme_classic() +
  theme(axis.text = element_blank(),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank()) ->theme_use2


## crosslink project
cl <- crosslink(nodes,edges,cross.by="type")

cl %>% cl_plot()

cl <- set_header(cl,header=unique(get_cross(cl)$cross))

cl %>% layout_polygon(crosses = c("Mir","Meth","Gene","Drug"),layout_based = "default") %>%
       tf_rotate(crosses= c("Mir","Meth","Gene","Drug"),angle = rep(45,4),layout="default") %>%
       tf_shift(x=0.2*(-1),y=1.5,crosses=c("Mir","Meth"),layout="default") -> cl


# plot annotation
top <- nodes$id[nodes$type == "Mir"] # set the order as you like
bottom <- nodes$id[nodes$type == "Gene"] # set the order as you like
right <- nodes$id[nodes$type == "Meth"] # set the order as you like

# Top plot
topAnn <-  mirData %>%
  mutate(mir_f = factor(mir, top)) %>%
  ggplot(mapping = aes(x=mir,y=-lfc)) +
  geom_bar(fill = "#E7298A",
           stat = "identity",
           width = 0.5) + # 间距跟bar同宽
  labs(x = NULL, y = "log2(Fold Change)") +
  theme_use2
topAnn

# Bottom plot
botAnn <- geneData %>%
  mutate(meth_f = factor(gene, bottom)) %>%
  ggplot(mapping = aes(x=gene,y=-lfc)) +
  geom_bar(fill =  RColorBrewer::brewer.pal(8, "Dark2")[c(1:5,7:9)][2],
           stat = "identity",
           width = 0.5) +
  labs(x = NULL, y = "Difference") +
  theme_use2
botAnn

# right plot
rgtAnn <- methData %>%
  mutate(mir_f = factor(meth, right)) %>%
  ggplot(mapping = aes(x=meth,y=-lfc)) +
  geom_bar(fill = RColorBrewer::brewer.pal(8, "Dark2")[c(1:5,7:9)][3],
           stat = "identity", 
           width = 0.5) +
  labs(x = NULL, y = "log2(Fold Change)") +
  theme_use2 +
  coord_flip()

rgtAnn

# left plot
mat = matrix(sample(1:100, 64, replace = T), nrow = 8)
colnames(mat)=as.character(cl@cross$Drug)
ggplot(data = melt(mat), aes(X1, X2, fill = value))+
     geom_tile(color = "white")+ 
     scale_fill_gradient2(low = "blue", high = "darkgreen", mid = "white",midpoint = 0) + 
     xlab("")+ylab("")+theme_classic()+
     theme(legend.position = "right",
          axis.line = element_blank(),
          axis.ticks = element_blank(),
           axis.text = element_blank()) -> lftAnno


cl_plot(cl,
       annotation=cl_annotation(top = topAnn,top.by = "Mir",top.height = 0.5,
                                bottom = botAnn,bottom.by = "Gene",bottom.height = 0.5,
                                right = rgtAnn,right.by ="Meth",right.width = 0.5,
                                left = lftAnno,left.by = "Drug" ,left.width = 0.5),
       cross = list( mapping = aes(color = type,size=degree,shape=type),
       scale = list(color = scale_color_manual(values = RColorBrewer::brewer.pal(8, "Dark2")[c(1:5,7:9)]),
       shape = scale_shape_manual(values = 15:23),
       size  = scale_size_continuous(range=c(1,5)))),
       link  = list(mapping = aes(color = type,linetype=type,size=cor),
                    scale = list(color = scale_color_manual(values = RColorBrewer::brewer.pal(8, "Set1")[c(1:5,7:9)]),
                                linetype=scale_linetype_manual(values = c(1:4)),
                                size=scale_size(range = c(0.5,1.5)))),
       header=NA,
       label = list(color="black"
                    ,angle=c(rep(0,8),rep(90,10),rep(0,10),rep(90,6))
                    ,nudge_x=c(rep(-2,8),rep(0,10),rep(2,10),rep(0,6))
                    ,nudge_y=c(rep(0,8),rep(-2,10),rep(0,10),rep(2,6))),
       add   =  theme(panel.background = element_blank(),axis.title = element_blank(),
                      panel.grid = element_blank(),
                      axis.ticks = element_blank(),
                      axis.text = element_blank()))


```
